Supplementary Material for Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-06-08 supp_mat_complexity_selection.Rmd
reorganizing analyses

Load dataset. This dataset is a subset that only includes larva that survive to pupation or turned into parasitoids. We omitted any data where larva mortality was unclear. This enabled us to focus on the selection pressures imposed by insect parasitoids rather than other sources.
```{r Preliminaries, message=FALSE, warning=FALSE, include=FALSE}

## LOAD REQUIRED LIBRARIES & SET OPTIONS ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gamm4) # generalized additive mixed-models
library(gsg) # calculating selection gradients
library(visreg)
library(viridis) # for color palette

## LOAD & MANAGE DATASET ----

@@

## CREATE SEPARATE DATASETS FOR SELECTION ANALYSES ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") %>%
  mutate(sc.gall_size = (Gall_Height_mm - mean(Gall_Height_mm))/sd(Gall_Height_mm),
         sc.clutch_size = (gall_individuals - mean(gall_individuals))/sd(gall_individuals),
         sc.female_preference = (Density_per_100_shoots - mean(Density_per_100_shoots))/sd(Density_per_100_shoots))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion") %>%
  mutate(sc.gall_size = (Gall_Height_mm - mean(Gall_Height_mm))/sd(Gall_Height_mm),
         sc.clutch_size = (gall_individuals - mean(gall_individuals))/sd(gall_individuals),
         sc.female_preference = (Density_per_100_shoots - mean(Density_per_100_shoots))/sd(Density_per_100_shoots))
```

Custom functions to avoid repetitive code
```{r FUNCTIONS}
gamm.model <- function(formula, data) {
  gamm4(formula, data, 
        random = ~(1|Genotype/Plant_Position/Gall_Number), 
        family = binomial(link = logit))
}

@@

Beta_avg_grad <- function(logistic_model, mean_fitness) {
  brackets <- predict(logistic_model, type = "response") * (1 - predict(logistic_model, type="response"))
  gradient <- mean(brackets) * fixef(logistic_model) / mean_fitness 
  return(gradient)
}
```

```{r}
# convert treatment to factor to use as 'by' variable in gamm.model
gall_selection.df <- gall_selection.df %>%
  mutate(Treatment_contr.sum=factor(Treatment.focus), 
         sc.gall_size = as.numeric(scale(Gall_Height_mm)),
         sc.clutch_size = as.numeric(scale(gall_individuals)),
         sc.female_preference = as.numeric(scale(Density_per_100_shoots)))
contrasts(gall_selection.df$Treatment_contr.sum) <- 'contr.sum'
# Inspect contrasts:
contrasts(gall_selection.df$Treatment_contr.sum)

## AIC MODEL SELECTION
# Note that I use gamm4 to fit all models to maintain consistency

## GAMM full
m1 <- gamm.model(gall_survival ~ Treatment_contr.sum +
                   s(sc.gall_size, by=Treatment_contr.sum) + 
                   s(sc.clutch_size, by=Treatment_contr.sum) + 
                   s(sc.female_preference, by=Treatment_contr.sum) + 
                   sc.gall_size:sc.clutch_size + sc.gall_size:sc.clutch_size:Treatment_contr.sum +
                   sc.gall_size:sc.female_preference + sc.gall_size:sc.female_preference:Treatment_contr.sum +                                 sc.clutch_size:sc.female_preference + sc.clutch_size:sc.female_preference:Treatment_contr.sum,
                 data = gall_selection.df)
summary(m1$gam); summary(m1$mer)

gamm_reduced <- gamm.model(gall_survival ~ Treatment_contr.sum +
                             s(sc.gall_size) + 
                             s(sc.clutch_size, by=Treatment_contr.sum) + 
                             s(sc.female_preference, by=Treatment_contr.sum),
                 data = gall_selection.df)
summary(gamm_reduced$gam); summary(gamm_reduced$mer)

gamm_reduced2 <- gamm.model(gall_survival ~ Treatment_contr.sum +
                             sc.gall_size + 
                             sc.clutch_size + sc.clutch_size:Treatment_contr.sum + 
                             s(sc.female_preference, by=Treatment_contr.sum),
                 data = gall_selection.df)
summary(gamm_reduced2$gam); summary(gamm_reduced2$mer)

GLMM_reduced <- gamm.model(gall_survival ~ Treatment_contr.sum +
                             sc.gall_size + 
                             sc.clutch_size + sc.clutch_size:Treatment_contr.sum + 
                             sc.female_preference + sc.female_preference:Treatment_contr.sum,
                            data = gall_selection.df)
summary(GLMM_reduced$gam); summary(GLMM_reduced$mer)

AIC(m1$mer, gamm_reduced$mer, gamm_reduced2$mer, GLMM_reduced$mer)

glmer_reduced <- glmer(gall_survival ~ Treatment_contr.sum +
                             sc.gall_size + 
                             sc.clutch_size + sc.clutch_size:Treatment_contr.sum + 
                             sc.female_preference + sc.female_preference:Treatment_contr.sum +
                             (1|Genotype/Plant_Position/Gall_Number),
                       family=binomial(link=logit),
                       data = gall_selection.df, glmerControl(optimizer="bobyqa"))
summary(glmer_reduced)
overdisp_fun(glmer_reduced) # no evidence of overdispersion

v <- visreg(fit, 'Days', 'Subject', re.form=(~Days|Subject), plot=FALSE)
plot(v)
plot(v, overlay=TRUE)

test <- visreg(glmer_reduced, xvar=c("sc.gall_size"), by="Genotype", re.form=(~1|Genotype), plot=FALSE, 
       scale="response", partial=TRUE, gg=TRUE, overlay=TRUE,
       cond = list(sc.gall_size=0, sc.clutch_size=0, sc.female_preference=0))
plot(test, partial=TRUE, overlay=TRUE)
visreg(glmer_reduced, xvar=c("sc.gall_size"), scale="response", partial=TRUE, gg=TRUE, overlay=TRUE)
visreg(glmer_reduced, xvar=c("sc.clutch_size"), by="Treatment_contr.sum", scale="response", partial=TRUE, gg=TRUE, overlay=TRUE)
visreg(glmer_reduced, xvar=c("sc.female_preference"), by="Treatment_contr.sum", scale="response", partial=TRUE, gg=TRUE, overlay=TRUE)

## GLMER full
glmer_full <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference)^2 +
                           Treatment_contr.sum*(I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2)), 
                         data = gall_selection.df)
summary(glmer_full$gam); summary(glmer_full$mer)
round(vcov(glmer_full$mer), 2) # no evidence of correlations that would confound interpretations of statistical main and interaction effects.

glmer_reduced <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference) +
                           I(sc.gall_size^2) + I(sc.female_preference^2) + sc.gall_size:sc.clutch_size, 
                         data = gall_selection.df)
summary(glmer_reduced$gam); summary(glmer_reduced$mer, correlation=TRUE)
visreg(glmer_reduced$gam, xvar="sc.gall_size", by="sc.clutch_size", scale="response")

glmer_reduced2 <- gamm.model(gall_survival ~ Treatment_contr.sum*sc.clutch_size + sc.gall_size + sc.female_preference +
                               I(sc.female_preference^2) + sc.gall_size:sc.clutch_size, 
                         data = gall_selection.df)
summary(glmer_reduced2$gam); summary(glmer_reduced2$mer)

AIC(glmer_full$mer, glmer_reduced$mer, glmer_reduced2$mer)

## LMER full
m3 <- gamm4(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference)^2 +
                   Treatment_contr.sum*(I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2)),
            family = gaussian(link=identity),
            random = ~(1|Genotype/Plant_Position/Gall_Number),
            data = gall_selection.df)

AIC(m1$mer, m2$mer, m3$mer) # suggests that the GLMM is best

## Do we need a model with interactions between traits?
m4 <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference) + #^2 +
                   Treatment_contr.sum*(I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2)), 
                 data = gall_selection.df)
AIC(m2$mer, m4$mer) # AIC suggests not

## Do we need a model with non-linear trait effects?
m5 <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference) + #^2 +
                   (I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2)), 
                 data = gall_selection.df)
AIC(m4$mer, m5$mer) # 
summary(m5$gam)
summary(m5$mer)

m6 <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference), 
                 data = gall_selection.df)
AIC(m5$mer, m6$mer) 
```

```{r}
## Calculate directional selection gradients
# following method of Janzen and Stern

Beta_avg_grad <- function(.) {
  mean_fitness <- inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness 
}

@@

```{r}

## Control treatment

# GLMM
glmer_control <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = control_df, glmerControl(optimizer="bobyqa"))
summary(glmer_control)

# Mean fitness
Wbar_control <- inverse_logit(fixef(update(glmer_control, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

# Use parametric bootstrapping to get confidence intervals for selection gradients
boot_Beta_control <- bootMer(glmer_control, FUN = Beta_avg_grad, nsim=200, parallel="multicore", ncpus=32)

data.frame(Beta_avg_grad = round(Beta_avg_grad(glmer_control),2),
           lowerCI_2.5 = round(apply(boot_Beta_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),2),
           upperCI_97.5 = round(apply(boot_Beta_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),2)) %>%
  knitr::kable(.)
```

```{r}

## Ectoparasitoid exclusion treatment

# GLMM
glmer_treatment <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference + 
                           (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = treatment_df, glmerControl(optimizer="bobyqa"))
summary(glmer_treatment)

# Mean fitness
Wbar_treatment <- inverse_logit(fixef(update(glmer_treatment, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

# Use parametric bootstrapping to get confidence intervals for selection gradients
boot_Beta_treatment <- bootMer(glmer_treatment, FUN = Beta_avg_grad, nsim=200, parallel="multicore", ncpus=32)

data.frame(Beta_avg_grad = round(Beta_avg_grad(glmer_treatment),2),
           lowerCI_2.5 = round(apply(boot_Beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),2),
           upperCI_97.5 = round(apply(boot_Beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),2)) %>%
  knitr::kable(.)
```

```{r}
## Custom function to calculate relative fitness intervals for my models
relative_fitness_intervals <- function(logistic_model, newdata, bootstraps){
  model_matrix <- model.matrix(~ sc.gall_size + sc.clutch_size + sc.female_preference, newdata)

  get_absolute_fitness <- function(.) {
  absolute_fitness <- inverse_logit(model_matrix %*% fixef(.))
  absolute_fitness 
  }

  get_bootstraps <-bootMer(logistic_model, FUN = get_absolute_fitness, nsim=bootstraps, parallel="multicore", ncpus=32)
  get_intervals <- apply(get_bootstraps$t, 2, function(x) x[order(x)][c(round(bootstraps*0.025,0),round(bootstraps*0.975,0))])

  mean_fitness <- mean(get_absolute_fitness(logistic_model))

  return(data.frame(relative_fitness = get_absolute_fitness(logistic_model)/mean_fitness,
                    lower_2.5 = get_intervals[1, ]/mean_fitness,
                    upper_97.5 = get_intervals[2, ]/mean_fitness,
                    newdata))
}
```

@@

```{r}
HPD_50_treatment_size <- coda::HPDinterval(coda::as.mcmc(treatment_df$sc.gall_size), prob=0.5)
HPD_50_control_size <- coda::HPDinterval(coda::as.mcmc(control_df$sc.gall_size), prob=0.5)

@@

RF_treatment_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = treatment_df$sc.gall_size, 
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=1000)

@@

RF_control_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = treatment_df$sc.gall_size,
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=1000)

@@

```{r}
RF_treatment_clutch_preference <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = distinct(treatment_df, Plant_Position, Gall_Number, sc.clutch_size)$sc.clutch_size,
     sc.female_preference = distinct(treatment_df, Plant_Position, sc.female_preference)$sc.female_preference), 
  bootstraps=100)

@@

```{r}
distribution_female_preference <- distinct(.data=gall_selection.df, Plant_Position, Density_per_100_shoots)$Density_per_100_shoots 
HPDinterval_female_preference <- coda::HPDinterval(coda::as.mcmc(scale(distribution_female_preference)), prob=0.95); HPDinterval_female_preference

>>>>

Load dataset. This dataset is a subset that only includes larva that survive to pupation or turned into parasitoids. We omitted any data where larva mortality was unclear. This enabled us to focus on the selection pressures imposed by insect parasitoids rather than other sources.
```{r setup}
knitr::opts_chunk$set(include = FALSE)

## Load required libraries ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gamm4) # generalized additive mixed-models
library(gsg) # calculating selection gradients
library(visreg)
library(viridis) # for color palette

## Load and manage data ----

@@

# convert treatment to factor to use as 'by' variable in generalized additive models
gall_selection.df <- gall_selection.df %>%
  mutate(Treatment_contr.sum=factor(Treatment.focus), 
         sc.gall_size = as.numeric(scale(Gall_Height_mm)),
         sc.clutch_size = as.numeric(scale(gall_individuals)),
         sc.female_preference = as.numeric(scale(Density_per_100_shoots)))
contrasts(gall_selection.df$Treatment_contr.sum) <- 'contr.sum'

## Create separate datasets for quantifying selection gradients ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") %>%
  mutate(sc.gall_size = (Gall_Height_mm - mean(Gall_Height_mm))/sd(Gall_Height_mm),
         sc.clutch_size = (gall_individuals - mean(gall_individuals))/sd(gall_individuals),
         sc.female_preference = (Density_per_100_shoots - mean(Density_per_100_shoots))/sd(Density_per_100_shoots))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion") %>%
  mutate(sc.gall_size = (Gall_Height_mm - mean(Gall_Height_mm))/sd(Gall_Height_mm),
         sc.clutch_size = (gall_individuals - mean(gall_individuals))/sd(gall_individuals),
         sc.female_preference = (Density_per_100_shoots - mean(Density_per_100_shoots))/sd(Density_per_100_shoots))

## Custom functions to avoid repetitive code ----
gamm.model <- function(formula, data) {
  gamm4(formula, data, 
        random = ~(1|Genotype/Plant_Position/Gall_Number), 
        family = binomial(link = logit))
}

@@

#Beta_avg_grad <- function(logistic_model, mean_fitness) {
#  brackets <- predict(logistic_model, type = "response") * (1 - predict(logistic_model, type="response"))
#  gradient <- mean(brackets) * fixef(logistic_model) / mean_fitness 
#  return(gradient)
#}

## Calculate directional selection gradients
# following method of Janzen and Stern 1998, Evolution
Beta_avg_grad <- function(.) {
  mean_fitness <- inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness 
}

## Calculate confidence intervals of relative fitness
relative_fitness_intervals <- function(logistic_model, newdata, bootstraps){
  model_matrix <- model.matrix(~ sc.gall_size + sc.clutch_size + sc.female_preference, newdata)

  get_absolute_fitness <- function(.) {
  absolute_fitness <- inverse_logit(model_matrix %*% fixef(.))
  absolute_fitness#/mean(absolute_fitness) 
  }

  get_bootstraps <-bootMer(logistic_model, FUN = get_absolute_fitness, nsim=bootstraps, parallel="multicore", ncpus=32)
  browser()

  get_intervals <- apply(get_bootstraps$t, 2, function(x) x[order(x)][c(round(bootstraps*0.025,0),round(bootstraps*0.975,0))])

  #mean_fitness <- mean(get_absolute_fitness(logistic_model))

  return(data.frame(relative_fitness = get_absolute_fitness(logistic_model),#/mean_fitness,
                    lower_2.5 = get_intervals[1, ],#/mean_fitness,
                    upper_97.5 = get_intervals[2, ],#/mean_fitness,
                    newdata))
}
```

We used Aikaike Information Criteria (AIC)

```{r AIC Model Selection}

## GAMM 1: models the effect of food-web treatment on univariate and multivariate fitness landscapes ----
GAMM_1 <- gamm.model(gall_survival ~ 
                       Treatment_contr.sum +
                       s(sc.gall_size, by=Treatment_contr.sum) + 
                       s(sc.clutch_size, by=Treatment_contr.sum) + 
                       s(sc.female_preference, by=Treatment_contr.sum) + 
                       sc.gall_size:sc.clutch_size + sc.gall_size:sc.clutch_size:Treatment_contr.sum +
                       sc.gall_size:sc.female_preference + sc.gall_size:sc.female_preference:Treatment_contr.sum +                                sc.clutch_size:sc.female_preference + sc.clutch_size:sc.female_preference:Treatment_contr.sum,
                     data = gall_selection.df)
# summary(GAMM_1$gam); summary(GAMM_1$mer)

## GAMM 2: models the effect of food-web treatment on univariate, but not multivariate landscape ----
GAMM_2 <- gamm.model(gall_survival ~ 
                       Treatment_contr.sum +
                       s(sc.gall_size, by=Treatment_contr.sum) + 
                       s(sc.clutch_size, by=Treatment_contr.sum) + 
                       s(sc.female_preference, by=Treatment_contr.sum) +
                       sc.gall_size:sc.clutch_size + 
                       sc.gall_size:sc.female_preference +                        
                       sc.clutch_size:sc.female_preference,
                     data = gall_selection.df)
# summary(GAMM_2$gam); summary(GAMM_2$mer)

## GAMM 3: models the effect of food-web treatment on univariate landscape ----
GAMM_3 <- gamm.model(gall_survival ~ 
                       Treatment_contr.sum +
                       s(sc.gall_size, by=Treatment_contr.sum) + 
                       s(sc.clutch_size, by=Treatment_contr.sum) + 
                       s(sc.female_preference, by=Treatment_contr.sum),
                      data = gall_selection.df)
# summary(GAMM_3$gam); summary(GAMM_3$mer)

## GAMM 4: assumes that the food-web does not modify the fitness landscape 
GAMM_4 <- gamm.model(gall_survival ~ 
                       Treatment_contr.sum +
                       s(sc.gall_size) + 
                       s(sc.clutch_size) + 
                       s(sc.female_preference),
                     data = gall_selection.df)
# summary(GAMM_4$gam); summary(GAMM_4$mer)

## AIC Comparison of GAMMs
knitr::kable(AIC(GAMM_1$mer, GAMM_2$mer, GAMM_3$mer, GAMM_4$mer))

@@

AIC model comparison suggests that ....

Now let's compare the best GAMMs to analog generalized linear mixed-models (GLMM). If these GLMMs have a lower AIC value than their respective GAMMs, then this suggests that there is only evidence of directional selection in our data.
```{r}
# Note that I still use gamm.model to allow for appropriate model comparisons

## GLMM 3: only models the effect of food-web treatment on directional selection gradients ----
GLMM_3 <- gamm.model(gall_survival ~ 
                       Treatment_contr.sum +
                       sc.gall_size + sc.gall_size:Treatment_contr.sum + 
                       sc.clutch_size + sc.clutch_size:Treatment_contr.sum + 
                       sc.female_preference + sc.female_preference:Treatment_contr.sum,
                      data = gall_selection.df)
#summary(GLMM_3$gam); summary(GLMM_3$mer)

## GLMM 4: only models the effect of food-web treatment on directional selection gradients ----
GLMM_4 <- gamm.model(gall_survival ~ 
                       Treatment_contr.sum +
                       sc.gall_size + 
                       sc.clutch_size + 
                       sc.female_preference,
                      data = gall_selection.df)
#summary(GLMM_4$gam); summary(GLMM_4$mer)

knitr::kable(AIC(GAMM_3$mer, GLMM_3$mer, GAMM_4$mer, GLMM_4$mer)) 
```

It appears that the best predictive model is the GLMM that includes an effect of food-web treatment on directional selection gradients. Therefore, we continue by quantifying the effect of food-web treatment on directional selection gradients.

The model selection procedure was useful for pairing down the complexity of our model. To actually quantify the directional selection gradients in each treatment though, we analyze each treatment separately.

```{r Selection Gradients - Complex Food Web}
# note that I've switched to using the lme4 package to analyzing the GLMM

# GLMM ----
glmer_control <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         data = control_df,
                         family = binomial(link = logit), optimizer = "bobyqa")
# summary(glmer_control)

# Mean fitness ----
Wbar_control <- inverse_logit(fixef(update(glmer_control, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

# Confidence intervals for selection gradients ----
boot_Beta_control <- bootMer(glmer_control, FUN = Beta_avg_grad, nsim=1000, parallel="multicore", ncpus=32)

## Summary ----
summary_Beta_glmm_control <- data.frame(
  Phenotype = c("Gall size", "Clutch size", "Female preference"),
  Food_Web = "Complex",
  Model = "GLMM",
  Beta = round(Beta_avg_grad(glmer_control),3),
  lower_2.5 = round(apply(boot_Beta_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),3),
  upper_97.5 = round(apply(boot_Beta_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),3)) 
```

```{r Selection Gradients - Simple Food Web}

## GLMM ----
glmer_treatment <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         data = treatment_df,
                         family = binomial(link = logit), optimizer = "bobyqa")
# summary(glmer_treatment)

## Mean fitness ----
Wbar_treatment <- inverse_logit(fixef(update(glmer_treatment, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

## Confidence intervals for selection gradients ----
boot_Beta_treatment <- bootMer(glmer_treatment, FUN = Beta_avg_grad, nsim=1000, parallel="multicore", ncpus=32)

## Summary ----
summary_Beta_glmm_treatment <- data.frame(
  Phenotype = c("Gall size", "Clutch size", "Female preference"),
  Food_Web = "Simple",
  Model = "GLMM",
  Beta = round(Beta_avg_grad(glmer_treatment),2),
  lower_2.5 = round(apply(boot_Beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),2),
  upper_97.5 = round(apply(boot_Beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),2)) 
```

To ensure we are calculating these selection gradients correctly, we'll compare our results to those obtained from the *gsg* package developed by Michael Morrissey.
```{r Check Gradient Calculations}

## GAMM Control ----
gamm_control <- gamm.model(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference,
                             data = control_df)
gsg_control <- data.frame(
  Phenotype = c("Gall size", "Clutch size", "Female preference"),
  Food_Web = "Complex",
  Model = "GAMM",
  Beta = round(c(gam.gradients(gamm_control$gam, 
                         phenotype = "sc.gall_size", covariates = c("sc.clutch_size","sc.female_preference"), 
                         se.method = "n")["B-sc.gall_size","estimates"],
           gam.gradients(gamm_control$gam, 
                         phenotype = "sc.clutch_size", covariates = c("sc.gall_size","sc.female_preference"), 
                         se.method = "n")["B-sc.clutch","estimates"],
           gam.gradients(gamm_control$gam, 
                         phenotype = "sc.female_preference", covariates = c("sc.gall_size","sc.clutch_size"),
                         se.method = "n")["B-sc.female_preference","estimates"]),3),
  lower_2.5 = NA,
  upper_97.5 = NA)

## GAMM Treatment ----
gamm_treatment <- gamm.model(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference,
                             data = treatment_df)
gsg_treatment <- data.frame(
  Phenotype = c("Gall size", "Clutch size", "Female preference"),
  Food_Web = "Simple",
  Model = "GAMM",
  Beta = round(c(gam.gradients(gamm_treatment$gam, 
                         phenotype = "sc.gall_size", covariates = c("sc.clutch_size","sc.female_preference"), 
                         se.method = "n")["B-sc.gall_size","estimates"],
           gam.gradients(gamm_treatment$gam, 
                         phenotype = "sc.clutch_size", covariates = c("sc.gall_size","sc.female_preference"), 
                         se.method = "n")["B-sc.clutch","estimates"],
           gam.gradients(gamm_treatment$gam, 
                         phenotype = "sc.female_preference", covariates = c("sc.gall_size","sc.clutch_size"),
                         se.method = "n")["B-sc.female_preference","estimates"]),3),
  lower_2.5 = NA,
  upper_97.5 = NA)

## Compare GAMM and GLMM estimates ----
compare_Beta <- bind_rows(gsg_control, gsg_treatment, summary_Beta_glmm_control, summary_Beta_glmm_treatment) %>%
  select(-lower_2.5, -upper_97.5) %>%
  spread(Model, Beta)

knitr::kable(compare_Beta)

The comparisons are quite close, but why should they differ at all? This suggests that my implementation of the Janzen and Stern 1998 method and Morrisey's are slightly different. Perhaps this is because Janzen and Stern relies on the coefficient to calculate the gradient, whereas Morrisey calculates it using the finite difference methods for partial derivatives...

Now let's look at how the confidence intervals vary between the GLMM and GAMMs. Note that I now fit it with a spline function to allow gsg to calculate bootstrap intervals (it won't otherwise).

```{r gsg Comparison}

## Fit generalized additive mixed models ----
gamm_control_2 <- gamm.model(gall_survival ~ s(sc.gall_size) + s(sc.clutch_size,k=9) + s(sc.female_preference),
                             data = control_df)

gamm_treatment_2 <- gamm.model(gall_survival ~ s(sc.gall_size) + s(sc.clutch_size,k=9) + s(sc.female_preference),
                             data = treatment_df)

## Estimate selection gradients ----
sg.size_control <- gradient.calc(mod = gamm_control_2$gam, 
                                 phenotype = "sc.gall_size", 
                                 covariates = c("sc.clutch_size","sc.female_preference"))
sg.size_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
                                   phenotype = "sc.gall_size", 
                                   covariates = c("sc.clutch_size","sc.female_preference"))
sg.clutch_control <- gradient.calc(mod = gamm_control_2$gam, 
                                        phenotype = "sc.clutch_size", 
                                        covariates = c("sc.gall_size","sc.female_preference"))
sg.clutch_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
                                          phenotype = "sc.clutch_size", 
                                          covariates = c("sc.gall_size","sc.female_preference"))
sg.preference_control <- gradient.calc(mod = gamm_control_2$gam, 
                                    phenotype = "sc.female_preference", 
                                    covariates = c("sc.clutch_size","sc.gall_size"))
sg.preference_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
                                      phenotype = "sc.female_preference", 
                                      covariates = c("sc.clutch_size","sc.gall_size"))

## Organize bootstrap estimates ----
bootstrap_df <- bind_rows(
  data.frame(Bootstrap_Estimates = sg.size_control$boot[ ,1], 
             Phenotype = "Gall size", Selection_Type = "Beta", Food_Web = "Complex"),
  data.frame(Bootstrap_Estimates = sg.size_treatment$boot[ ,1], 
             Phenotype = "Gall size", Selection_Type = "Beta", Food_Web = "Simple"),
  data.frame(Bootstrap_Estimates = sg.clutch_control$boot[ ,1], 
             Phenotype = "Clutch size", Selection_Type = "Beta", Food_Web = "Complex"),
  data.frame(Bootstrap_Estimates = sg.clutch_treatment$boot[ ,1], 
             Phenotype = "Clutch size", Selection_Type = "Beta", Food_Web = "Simple"),
  data.frame(Bootstrap_Estimates = sg.preference_control$boot[ ,1], 
             Phenotype = "Female preference", Selection_Type = "Beta", Food_Web = "Complex"),
  data.frame(Bootstrap_Estimates = sg.preference_treatment$boot[ ,1], 
             Phenotype = "Female preference", Selection_Type = "Beta", Food_Web = "Simple"))

## Summarise bootstrap estimates ----
gsg_gradients <- bootstrap_df %>%
  group_by(Phenotype, Food_Web) %>%
  summarise(Beta = round(mean(Bootstrap_Estimates),3),
            lower_2.5 = round(quantile(Bootstrap_Estimates, probs = 0.025),3),
            upper_97.5 = round(quantile(Bootstrap_Estimates, probs = 0.975),3)) %>%
  mutate(Model = "gsg_GAMM")

## Compare gsg GAMM and GLMM estimates of directional selection ----
bind_rows(gsg_gradients, summary_Beta_glmm_control, summary_Beta_glmm_treatment) %>%
  arrange(Phenotype, Food_Web, Model) %>%
  mutate(absolute_range = upper_97.5 - lower_2.5) %>%
  knitr::kable(.)
```

Both methods give the same qualitative results. However, range of the confidence intervals can be more than twice as large using my GLMM bootstrapping methods. Therefore, our reported intervals provide a conservative estimate.

```{r Compare Selection Gradients across Treatments - GLMM}

## Gather bootstrap data from GLMMs
glmm_boostrap_df <- bind_rows(
  data.frame(boot_Beta_treatment$t, ID = rep(1:dim(boot_Beta_treatment$t)[1]), Food_Web = "Simple"),
  data.frame(boot_Beta_control$t, ID = rep(1:dim(boot_Beta_control$t)[1]), Food_Web = "Complex")
)

## Gall size ----
select(glmm_boostrap_df, sc.gall_size, ID, Food_Web) %>%
  spread(Food_Web, sc.gall_size) %>%
  mutate(diff_gradient = Complex - Simple) %>%
  summarise(p_value = sum(diff_gradient < 0)/dim(glmm_boostrap_df)[1])

## Clutch size ----
select(glmm_boostrap_df, sc.clutch_size, ID, Food_Web) %>%
  spread(Food_Web, sc.clutch_size) %>%
  mutate(diff_gradient = Complex - Simple) %>%
  summarise(p_value = sum(diff_gradient < 0)/dim(glmm_boostrap_df)[1])

## Female preference ----
select(glmm_boostrap_df, sc.female_preference, ID, Food_Web) %>%
  spread(Food_Web, sc.female_preference) %>%
  mutate(diff_gradient = Complex - Simple) %>%
  summarise(p_value = sum(diff_gradient < 0)/dim(glmm_boostrap_df)[1])
```

We observe the same qualitative results with *gsg* GAMMs, except that the P-values are always lower.

```{r Compare Selection Gradients across Treatments - gsg GAMM}

## Gall size ----
bootstrap_df %>%
  filter(Phenotype == "Gall size", Selection_Type == "Beta") %>%
  mutate(ID = rep(1:1000,2)) %>%
  spread(Food_Web, Bootstrap_Estimates) %>%
  mutate(diff_gradient = Complex - Simple) %>%
  summarise(p_value = sum(diff_gradient < 0)/1000)

## Clutch size ----
bootstrap_df %>%
  filter(Phenotype == "Clutch size", Selection_Type == "Beta") %>%
  mutate(ID = rep(1:1000,2)) %>%
  spread(Food_Web, Bootstrap_Estimates) %>%
  mutate(diff_gradient = Complex - Simple) %>%
  summarise(p_value = sum(diff_gradient < 0)/1000)

## Female preference ----
bootstrap_df %>%
  filter(Phenotype == "Female preference", Selection_Type == "Beta") %>%
  mutate(ID = rep(1:1000,2)) %>%
  spread(Food_Web, Bootstrap_Estimates) %>%
  mutate(diff_gradient = Complex - Simple) %>%
  summarise(p_value = sum(diff_gradient < 0)/1000)
```

@@

```{r GLMM Univariate Landscapes, eval=FALSE}

## Gall size ----
HPD_50_treatment_size <- coda::HPDinterval(coda::as.mcmc(treatment_df$sc.gall_size), prob=0.5)
HPD_50_control_size <- coda::HPDinterval(coda::as.mcmc(control_df$sc.gall_size), prob=0.5)

@@

RF_treatment_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = treatment_df$sc.gall_size, 
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=10)

@@

RF_control_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = treatment_df$sc.gall_size,
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=100)

@@

```{r GLMM Multivariate Landscapes, eval=FALSE}

## Clutch size x Female Preference ----
RF_treatment_clutch_preference <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = distinct(treatment_df, Plant_Position, Gall_Number, sc.clutch_size)$sc.clutch_size,
     sc.female_preference = distinct(treatment_df, Plant_Position, sc.female_preference)$sc.female_preference), 
  bootstraps=100)

@@
